Parsing Postfix log files

### 2009-02-24 doc/logparser-implementation.tex
Annotated all URLs with last checked date.
Added delay, delays, and dsn to the results table; they're automatically     extracted and saved because filter_regex() fiddles the existing regex     snippets. Updated the thesis with new result columns, including the comparisons in the     state of the art review. Changed how the acronyms glossary is formatted by adding a new style that uses     eqlist. A couple of minor fixes.

XXX UPDATE REFERENCES WHEN ARCHITECTURE AND IMPLEMENTATION ARE MORE FLESHED
OUT\@.

@@

XXX WHEN FINISHED, CHECK ALL REFERENCES TO ENSURE THEY POINT AT THE CORRECT
CHAPTER\@.

@@

XXX LAST CHECKED DATES FOR ALL URLS\@.

@@

    \item \acronym{SQL} is reasonably standard and many people will already
        be familiar with it; for those unfamiliar with it there are lots of
        readily available resources from which to learn (a good
        introduction to \acronym{SQL} can be found at
        \url{http://philip.greenspun.com/sql/}, others are
        \url{http://www.w3schools.com/sql/default.asp},
        \url{http://sqlcourse.com/}).  Although every vendor implements a
        different dialect of \acronym{SQL}, the basics are the same
        everywhere (analogous to the overall similarities and minor
        differences amongst Irish English, British English, American
        English, and Australian English).  Depending on the database in use
        there may be tools available that reduce or remove the requirement
        to know \acronym{SQL}; several are available for \gls{SQLite}, the
        default database used by the parser implementation.

@@

@@

@@

\begin{verbatim}
NOQUEUE: reject: RCPT from example.com[192.0.2.1]: 550
  <foo@example.com>: Sender address rejected: Domain not found;
  from=<foo@example.com> to=<info@example.net>
  proto=SMTP helo=<smtp.example.com>
\end{verbatim}

@@

SEE TABLE (PROBABLY ON NEXT PAGE): THERE IS NEARLY A 1--1 MAPPING BETWEEN
action AND postfix\_action.  NEW ACTIONS COULD BE ADDED TO FORCE A 1--1
MAPPING, E.G. COMMIT + DISCARDED COULD BECOME A DISCARD ACTION\@.
MAIL\_QUEUED HAVING INFO AND PROCESSING CAN PROBABLY BE DEALT WITH, THE
BIGGEST DIFFICULTY WILL BE WITH SAVE\_DATA\@.

>>>>

XXX UPDATE REFERENCES WHEN ARCHITECTURE AND IMPLEMENTATION ARE MORE FLESHED
OUT\@.  XXX WHEN FINISHED, CHECK ALL REFERENCES TO ENSURE THEY POINT AT THE
CORRECT CHAPTER\@.

@@

@@

@@

    \item \acronym{SQL} is reasonably standard and many people will already
        be familiar with it; for those unfamiliar with it there are lots of
        readily available resources from which to learn (a good
        introduction to \acronym{SQL} can be found at
        \urlLastChecked{http://philip.greenspun.com/sql/}{2009/02/23},
        others are
        \urlLastChecked{http://www.w3schools.com/sql/default.asp}{2009/02/23},
        \urlLastChecked{http://sqlcourse.com/}{2009/02/23}).  Although
        every vendor implements a different dialect of \acronym{SQL}, the
        basics are the same everywhere (analogous to the overall
        similarities and minor differences amongst Irish English, British
        English, American English, and Australian English).  Depending on
        the database in use there may be tools available that reduce or
        remove the requirement to know \acronym{SQL}; several are available
        for \gls{SQLite}, the default database used by the parser
        implementation.

@@

    \item [dsn] The \acronym{DSN} code, similar to the \acronym{SMTP} code;
        it will only be present for log lines indicating delivery or
        rejection.  \acronym{DSN} support was added to Postfix in version
        2.3; logs generated by previous versions will not contain any DSN
        information.

@@

    \item [delay] How long the mail was delayed for while it was being
        delivered.  This will only be present for delivered mails.

    \item [delays] More details information about how long the mail was
        delayed for while it was being delivered  This will only be present
        for delivered mails..

@@

\begin{verbatim}
NOQUEUE: reject: RCPT from example.com[192.0.2.1]: 550 5.1.8
  <foo@example.com>: Sender address rejected: Domain not found;
  from=<foo@example.com> to=<info@example.net>
  proto=SMTP helo=<smtp.example.com>
\end{verbatim}

@@

XXX SEE TABLE BELOW\@: THERE IS NEARLY A 1--1 MAPPING BETWEEN action AND
postfix\_action.  NEW ACTIONS COULD BE ADDED TO FORCE A 1--1 MAPPING, E.G.
COMMIT + DISCARDED COULD BECOME A DISCARD ACTION\@.  MAIL\_QUEUED HAVING
INFO AND PROCESSING CAN PROBABLY BE DEALT WITH, THE BIGGEST DIFFICULTY WILL
BE WITH SAVE\_DATA\@.

### 2009-02-24 doc/logparser-other-parsers-review.tex
Added delay, delays, and dsn to the results table; they're automatically     extracted and saved because filter_regex() fiddles the existing regex     snippets. Updated the thesis with new result columns, including the comparisons in the     state of the art review. Changed how the acronyms glossary is formatted by adding a new style that uses     eqlist. A couple of minor fixes.

The same ten parsers have been reviewed and compared with \parsername{},
this project's finished parser, to show how much effort would have been
required to fulfil the aims and requirements of this project.  It is
important to compare and contrast newly developed algorithms and parsers
against those already available, to accurately judge what improvements, if
any, are delivered by the newcomers.  Each of the reviewed parsers was
tested with the \numberOFlogFILES{} test log files described in
\sectionref{parser efficiency}.  \Tableref{Summary of parsers' features}
summarises the results.

@@

    \item Only \parsername{} enables parsing of new log lines without
        extensive and intrusive modifications to the parser; \parsernames{}
        architecture is described in \sectionref{why separate rules,
        actions, and framework?}.

@@

The data extracted by \parsername{} is documented in
\sectionref{connections table} and \sectionref{results table}; for
convenience that list is repeated here: client and server \acronym{IP}
address and hostname, HELO hostname, queueid, start time, end time,
\acronym{SMTP} code, sender, recipient, size, message ID\@.

@@

Whether a MySQL database or Berkeley DB table is chosen in addition to the
\acronym{CSV} output, \acronym{LMA} stores the following data: time and
date, client hostname and \acronym{IP} address, server \acronym{IP}
address, sender and recipient addresses, \acronym{SMTP} code, and size (for
accepted mails only).  It is unclear what time and date is stored: start
time, end time, or delivery time?  Unlike \parsername{} it does not store
the server hostname, HELO hostname, queueid, start and end times,
timestamps for each log line, or message id (for accepted mails only).
Handling of multiple recipients, \acronym{SMTP} codes, or remote
servers\footnote{A single mail may be sent to multiple remote servers if it
was addressed to recipients in different domains, or Postfix needs to try
multiple servers for one or more recipients.} is not explained;
experimental observation shows that multiple records are added when a mail
had multiple recipients (sadly the records are not associated or linked in
any way), and presumably the same approach is taken when there were
multiple destination servers.

@@

    \item Some records output by DAILY-REJECT are truncated: they are
        missing the last | separating fields and the newline following it,
        so the line containing only a single space is concatenated with the
        record.

@@

Even if the three data sets were combined Sawmill would extract less data
than \parsername{}: it omits the HELO hostname, queueid, and start and end
times.  Sawmill does not extract any data about rejections except when the
rejection is caused by a \acronym{DNSBL} check (\texttt{RBL list} in the
list of fields).

@@

When parsing Postfix log files Splunk parses the standard
syslog\glsadd{syslog} fields at the beginning of the log line, and extracts
any \texttt{key=value} pairs occurring after the standard syslog prologue:
to and from addresses, HELO hostname, and protocol (\acronym{SMTP} or
\acronym{ESMTP}).  \parsername{} extracts noticeably more data (client and
server \acronym{IP} address and hostname, queueid, start and end times,
timestamps for each log line, \acronym{SMTP} code, and message ID), though
it does not make the full text of the line available (this could be
trivially added if desired, but would greatly increase the size of the
resulting database).  The full power of \acronym{SQL} is available when
searching the data extracted by \parsername{}, allowing the user to search
on arbitrarily complicated conditions.

@@

\noindent{}YAALA's Postfix parser extracts some of the fields \parsername{}
does: for client and server it stores either the \acronym{IP} address or
the hostname, not both; it omits the HELO hostname, queueid, \acronym{SMTP}
code, size of each accepted mail, start and end times, timestamps for each
log line, and message ID\@.  It extracts some data that \parsername{} does
not: the delay in delivering each mail, and how many times delivery was
deferred for each mail; these could easily be extracted by \parsername{} if
desired.  Unlike \parsername{}, YAALA does not maintain separate counters
for each restriction; this rules out the possibility of using the collected
data for optimisation, testing, or understanding of restrictions.  YAALA's
Postfix parser does not deal with the complications explained in
\sectionref{complications}, except it does correlate log lines by queueid.

@@

\noindent{}Examining the source code reveals that the parser looks for
\texttt{<key>=<value>} pairs in each log line, extracts them, and
correlates the data by queueid.  This approach will extract the following
data: HELO hostname, queueid, \acronym{SMTP} code, sender and recipient
addresses, and size of accepted mails.  It is unclear if the parser will
extract any further data.  Lire misses the following fields extracted by
\parsername{}: client and server \acronym{IP} address and hostname, start
and end times, timestamps of each log line, and message ID\@.

@@

It also counts the number of log lines parsed and skipped.  It omits client
\acronym{IP} address and hostname, server \acronym{IP} address, HELO
hostname, queueid, timestamps of each log line, and message ID\@.  It
extracts the delay for delivered mails, which \parsername{} does not.  Log
lines are correlated based on the queueid (called sessionname [sic] within
Logrep), but not by \acronym{pid}.  The parsing is error prone: empty
fields are saved when the log line does not match the regex, though it
appears that they will not overwrite existing data.  Most notably
rejections are completely ignored, making it unsuitable for the purposes of
this project.  It does not try to address any of the complications in
\sectionref{complications} except for correlating by queueid.

>>>>

The same ten parsers have been reviewed and compared with \parsername{},
this project's finished parser, to show how much effort would have been
required to fulfil the aims and requirements of this project.  It is
important to compare and contrast newly developed algorithms and parsers
against those already available, to accurately judge what improvements, if
any, are delivered by the newcomers.  Each of the reviewed parsers was
tested with the \numberOFlogFILES{} test log files described in
\sectionref{parser efficiency}.  \Tableref{Summary of parsers' features}
summarises the results of this review.

@@

    \item Only \parsername{} enables parsing of new log lines without
        extensive and intrusive modifications to the parser; \parsernames{}
        architecture is described in \sectionref{parser architecture}.

@@

The data extracted by \parsername{} is documented in
\sectionref{connections table} and \sectionref{results table}; for
convenience that list is repeated here: client and server \acronym{IP}
address and hostname, HELO hostname, queueid, start time, end time,
\acronym{SMTP} code, \acronym{DSN} code, sender, recipient, size, message
ID, delay and delays.

@@

Whether a MySQL database or Berkeley DB table is chosen in addition to the
\acronym{CSV} output, \acronym{LMA} stores the following data: time and
date, client hostname and \acronym{IP} address, server \acronym{IP}
address, sender and recipient addresses, \acronym{SMTP} code, and size (for
accepted mails only).  It is unclear what time and date is stored: start
time, end time, or delivery time?  Unlike \parsername{} it does not store
the server hostname, HELO hostname, queueid, start and end times,
timestamps for each log line, \acronym{DSN} code, delivery delays, or
message id (for accepted mails only).  Handling of multiple recipients,
\acronym{SMTP} codes, or remote servers\footnote{A single mail may be sent
to multiple remote servers if it was addressed to recipients in different
domains, or Postfix needs to try multiple servers for one or more
recipients.} is not explained; experimental observation shows that multiple
records are added when a mail had multiple recipients (sadly the records
are not associated or linked in any way), and presumably the same approach
is taken when there were multiple destination servers.

@@

    \item Some records output by DAILY-REJECT are truncated: they are
        missing the last $|$ that separates the fields and the newline
        following it, so the line containing only a single space is
        concatenated with the record.

@@

Even if the three data sets were combined Sawmill would extract less data
than \parsername{}: it omits the HELO hostname, queueid, \acronym{DSN}
code, delivery delays, and start and end times.  Sawmill does not extract
any data about rejections except when the rejection is caused by a
\acronym{DNSBL} check (\texttt{RBL list} in the list of fields).

@@

When parsing Postfix log files Splunk parses the standard
syslog\glsadd{syslog} fields at the beginning of the log line, and extracts
any \texttt{key=value} pairs occurring after the standard syslog prologue:
to and from addresses, HELO hostname, and protocol (\acronym{SMTP} or
\acronym{ESMTP}).  \parsername{} extracts noticeably more data (client and
server \acronym{IP} address and hostname, queueid, start and end times,
timestamps for each log line, \acronym{SMTP} and \acronym{DSN} codes,
delivery delays, and message ID), though it does not make the full text of
the line available (this could be trivially added if desired, but would
greatly increase the size of the resulting database).  The full power of
\acronym{SQL} is available when searching the data extracted by
\parsername{}, allowing the user to search on arbitrarily complicated
conditions.

@@

\noindent{}YAALA's Postfix parser extracts some of the fields \parsername{}
does: for client and server it stores either the \acronym{IP} address or
the hostname, not both; it omits the HELO hostname, queueid, \acronym{SMTP}
and \acronym{DSN} codes, size of each accepted mail, start and end times,
timestamps for each log line, and message ID\@.  It extracts some data that
\parsername{} does not: how many times delivery was deferred for each mail;
this information can be determined from the database populated by
\parsername{} if desired.  Unlike \parsername{}, YAALA does not maintain
separate counters for each restriction; this rules out the possibility of
using the collected data for optimisation, testing, or understanding of
restrictions.  YAALA's Postfix parser does not deal with the complications
explained in \sectionref{complications}, except it does correlate log lines
by queueid.

@@

\noindent{}Examining the source code reveals that the parser looks for
\texttt{<key>=<value>} pairs in each log line, extracts them, and
correlates the data by queueid.  This approach will extract the following
data: HELO hostname, queueid, \acronym{SMTP} code, sender and recipient
addresses, and size of accepted mails.  It is unclear if the parser will
extract any further data.  Lire misses the following fields extracted by
\parsername{}: client and server \acronym{IP} address and hostname, start
and end times, \acronym{DSN} code, delivery delays, timestamps of each log
line, and message ID\@.

@@

It also counts the number of log lines parsed and skipped.  It omits client
\acronym{IP} address and hostname, server \acronym{IP} address, HELO
hostname, queueid, timestamps of each log line, \acronym{DSN} code, and
message ID\@.  Log lines are correlated based on the queueid (called
sessionname [sic] within Logrep), but not by \acronym{pid}.  The parsing is
error prone: empty fields are saved when the log line does not match the
regex, though it appears that they will not overwrite existing data.  Most
notably rejections are completely ignored, making it unsuitable for the
purposes of this project.  It does not try to address any of the
complications in \sectionref{complications} except for correlating by
queueid.
